### 2013-04-24 paper/paper.tex
Converted the outline into a latex document and pdf that serve as a starting point for this writeup.

\documentclass[]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

\author{Joshua Branchaud \and Eric Rizzi}
\date{}

\begin{document}

\title{Jager: Symbolic Specification-Based Program Adjustment}

\maketitle

\section{Outline}

\subsection{Introduction}

\subsubsection{Background}

\subsubsection{Motivation}

\subsection{Approach}

\subsubsection{Trace Generation}

\subsubsection{Trace Differentiation}

\subsubsection{Fault Localization}

\subsubsection{Pre and Post Condition Generation}

\subsubsection{Synthesis}

\subsection{Evaluation}

\subsection{Related Work}

\subsection{Future Work}

\subsection{Conclusion}

\subsection{References}

\end{document}